---
title: "Practical Machine Learning - Weight Lifting Exercises"
author: "Chong Wen Mau"
date: "December 26, 2015"
output: html_document
---

##Overview
This report documents the prediction model building to predict the manner in which people did the exercise. 

The data from accelerometers on the belt, forearm, arm, and dumbell of 6 participant will be used as training data. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The training data consists of 160 variables and 19622 observations. The outcome variable is "classe".

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(AppliedPredictiveModeling)
library(caret)
library(rpart)

training = read.csv("pml-training.csv")
testing = read.csv("pml-testing.csv")
```

## Data Cleaning

1. In the data cleaning process, the computed variables, such as average, stanrdard deviation, minimum value, max value, variance, skewness, kurtosis and amplitude have been removed. 

2. Then, the data rows with missing values have been excluded for the model training.

```{r}

computedVarPattern = "avg|stddev|min|max|var|skewness|kurtosis|amplitude"
new_training = training[,names(training)[-grep(computedVarPattern, names(training))]]

new_training = new_training[complete.cases(new_training),]

#All measurement variables are converted to numeric type.

for (i in 7:59)
{
  new_training[,i] <- as.numeric(new_training[,i])
}
```

## Model Training

The model is trained using Classification and Regression Tree algorithm with 10- fold cross validation.

The classification tree algorithm is employed in model training because the outcome variable is categorical and the result is easier to be interpreted.

75 % of the training data are used for model training and the remaining are used for model testing.

```{r}

set.seed(12345)

inTrain = createDataPartition(new_training$classe, p = 0.75)[[1]]
trainSet = new_training[inTrain,]
testSet = new_training[-inTrain,]

trainControl = trainControl(method = "cv", number = 10)

#Include the measurement variables only
modFit = train(classe ~ ., method="rpart", trControl = trainControl, data = trainSet[-1:-7])

modFit$finalModel
```

##Model Evaluation

In sample error

```{r}
sum(predict(modFit, trainSet) != trainSet$classe) / nrow(trainSet)
```

Out of sample error
```{r}
sum(predict(modFit, testSet) != testSet$classe) / nrow(testSet)
```

